On the application of microwave acoustic resonant viral inactivation \thanks{I would be delighted to include any criticisms or comments anyone may have; preferably leave them on the GitHub issues page, or email therobotist@gmail.com, @0xDBFB7 on Twitter, or irc.0xDBFB7.com:6667 \#covid.}

### 2021-01-17 documents/background.tex
Progress

@@

HFM-189 found no published and replicated adverse health effects or biological mechanisms, beyond
thermal interaction, for pulses shorter than 100 ms which suggested that neither the peak E-field limit in the
IEEE C95.1TM-2005 safety standard nor the proposed limit in the Directive 2004/40/EC [33] (subsequently
promulgated as 2013/35/EU [34]) have scientific basis.

The group noted that physical laws governing the propagation of E-fields in air already limit the maximum allowable peak E-field at ~3 MV/m (air breakdown). Current research efforts by members to expose biological organism(s), tissues, and cells to environmental
fields up to this magnitude have been unable to elicit an acute biological response.

\footnote{``{[B]elow about 6 GHz, where EMFs penetrate deep into tissue (and thus require depth to be considered), it is useful to describe this in terms of “specific energy absorption rate” (SAR), which is the power absorbed per unit mass $(W/kg)$. Conversely, above 6 GHz, where EMFs are absorbed more superficially (making depth less relevant), it is useful to describe exposure in terms of the density of absorbed power over area $W/m^2$, which we refer to as “absorbed power density”}'' [ICNIRP 2020 \faExternalLink] }

@@

Pakhomov \cite{Comparative} expose frog heart cells. No effects.

@@

We also would like to point out that these normal modes
are damped when the virus particle is embedded inside a
liquid...when the virus is inside water, for example, there is not much acoustic impedance mismatch at the surface. For this reason the normal modes will be broad and have short lifetimes. Therefore, the objective of killing viruses by sending out sound waves that resonate and destroy them is probably unworkable in such a configuration.

@@

@@

Ultrafast femtosecond laser pulses

@@

Notice that a non-zero resonant frequency
requires
ωR 6= 0 ⇒ β < ω/√
2. (49)
That is, the oscillator must be sufficiently under-damped in order to exhibit a
resonance response

@@

>>>>

I was hoping I would have a definitive demonstration, so as to not redirect resources from more promising avenues.

TRL 0

(more data exists for influenza A; it is sufficiently structurally similar that ).

We don't want a purely membrane-electroporative effect since the lipid membrane of the virus is similar to (derived from) the host cell and little selectivity can be expected. Unless the virus has a lower capacitance and charges faster than the cell.

Also, the pore must stay open for long enough that damage is dealt to the virus.

Viral electroporation specifically has been observed.

Mizuno \cite{Inactivation1990} 100\% inactivate SVD Enterovirus (a 27 nm non-enveloped encapsidated ssRNA) and EHV (a 100 nm enveloped dsDNA) with 120 pulses, 500 or 600 nanoseconds each, at 3 MV/m.

The thermal artifact  the energy \footnote{(43 cal/cc * (3 microliters)) / (3 milligrams * 4.2 J/(g K))}. The electrodes were un-insulated, so electrochemical effects are possible.

Temperature rise

at a slow field of approximately $10^7$ V/m, in the Vaccinia poxvirus (a 300 nm enveloped dsDNA virus). \cite{AC2017}. Interestingly, the same effect is not observed in T4 \cite{Manipulation2013}.

\cite{comparative2003} discuss the relative inactivation of bacterial hosts and MS2 and PRD1 phages.

Electric field has many more free variables than other treatments;

best practices were not followed.

A wave propagating into tissue with $e_r=70$ has a refractive index $n=\sqrt{\mu/\epsilon}$, and is transmitted as $\frac{2 n_1}{n_1 + n_2}$; therefore about 0.2x the electric field is transmitted. This is of no concern for waveguide or bolus applicators, since the reflected energy is captured and returned.

@@

HFM-189 found no published and replicated adverse health effects or biological mechanisms, beyond
thermal interaction, for pulses shorter than 100 ms which suggested that neither the peak E-field limit in the
IEEE C95.1-2005 safety standard [n.b. 4.3, Table 9, subsection e. peak E field 100 kV/m] nor the proposed limit in the Directive 2004/40/EC [33] (subsequently promulgated as 2013/35/EU [34]) have scientific basis.

" physical laws governing the propagation of E-fields in air already limit the maximum allowable peak E-field at ~3 MV/m (air breakdown). Current research efforts by members to expose biological organism(s), tissues, and cells to environmental
fields up to this magnitude have been unable to elicit an acute biological response."

\footnote{A note on units, from ICNIRP: ``{[B]elow about 6 GHz, where EMFs penetrate deep into tissue (and thus require depth to be considered), it is useful to describe this in terms of “specific energy absorption rate” (SAR), which is the power absorbed per unit mass $(W/kg)$. Conversely, above 6 GHz, where EMFs are absorbed more superficially (making depth less relevant), it is useful to describe exposure in terms of the density of absorbed power over area $W/m^2$, which we refer to as “absorbed power density”}''. }

A provisional limit of 100 kV/m; the standard now that the fourier transform instantaneous pulse energy must not exceed 1/5 the energy over some specified averaging time.

C95.1-2005

When the waveform of the external electric field is non-sinusoidal, such as with pulsed or mixed frequency
waveforms, the rms value of the spatially averaged external field shall conform to the MPEs of Table 4, and
also to either of the criteria stated in 4.1.2.4.1 and 4.1.2.4.2.

@@

Pakhomov \cite{Comparative} expose frog heart cells. No effects on pacemaker at 0.9 kV/m (although the field inside the ).

Note: Not being biologists or microwave engineers, the trained reader may find that we are too verbose in simple technique and not in another.

\begin{fquote}[Saviot][ \cite{Comment2004}]
	We also would like to point out that these normal modes are damped when the virus particle is embedded inside a liquid...when the virus is inside water, for example, there is not much acoustic impedance mismatch at the surface. For this reason the normal modes will be broad and have short lifetimes. Therefore, the objective of killing viruses by sending out sound waves that resonate and destroy them is probably unworkable in such a configuration.	
\end{fquote}

@@

Perilla \cite{}, in a trustworthy explicitly solvated all-atom MD simulation, describe a few surface-wave modes.

electropermeabilization

@@

add "timeline"

@@

@@

Notice that a non-zero resonant frequency
requires $\beta = b/(2m) \i-t \sqrt{2}$ \cite{Driven}

@@

Matzumoto \cite{Inactivation1991} find that in one case, an arc was struck within the medium, and that the UV light produced was sufficient to inactivate the organism.

### 2021-01-17 documents/paper.tex
Progress

Industry Canada limits on the instantaneous electric field anywhere inside the body are $1.35 \times 10^-4$ V/m * Hz ()\cite{RSS1022015}, Table 4), $0.81 \times 10^6$ V / m \@ 6 GHz to $1.2 \times 10^6$ / m \@ 9 GHz

@@

>>>>

Industry Canada only appeaars to offer an instantaneous limit at $1.35 \times 10^-4$.

@@

\subfile{viscosity}
